\section*{Appendix B\\Recursive Lagrangian Structure and Field Dynamics}
\addcontentsline{toc}{section}{Appendix B: Recursive Lagrangian Structure and Field Dynamics}
\label{appendix:B}

\subsection*{B.1 Total Action Decomposition}

The recursive action per cycle is defined as:
\[
\mathcal{A}_n[\phi] = \int_{\Sigma_n} \left( 
\mathcal{L}_{\text{EH}} + 
\mathcal{L}_{\text{YM}} + 
\mathcal{L}_{\text{scalar}} + 
\mathcal{L}_{\text{int}} + 
\mathcal{L}_{\text{obs}} + 
\mathcal{L}_{\text{ent}} 
\right) \, d^4x,
\]
where each term corresponds to a physical component within the recursive structure:

\begin{itemize}[leftmargin=1.5em]
  \item $\mathcal{L}_{\text{EH}}$: Einstein–Hilbert term for gravitational curvature:
  \[
  \mathcal{L}_{\text{EH}} = \frac{1}{2\kappa} R,
  \quad \text{with } \kappa = 8\pi G.
  \]
  
  \item $\mathcal{L}_{\text{YM}}$: Gauge field term with coherence-modulated couplings:
  \[
  \mathcal{L}_{\text{YM}} = -\frac{1}{4} \sum_a \left( g_n^{(a)} \right)^{-2} F_{\mu\nu}^{(a)} F^{\mu\nu}_{(a)},
  \]
  where $g_n^{(a)} = f^{(a)}(\lambda_n, E_n, \gamma_n)$ reflects cycle-dependent coherence structure.

  \item $\mathcal{L}_{\text{scalar}}$: Scalar field evolution with attractor-modulated potential:
  \[
  \mathcal{L}_{\text{scalar}} = \frac{1}{2} g^{\mu\nu} \partial_\mu \varphi \partial_\nu \varphi - V(\varphi, \lambda_n).
  \]

  \item $\mathcal{L}_{\text{int}}$: Interaction term coupling scalar fields to internal symmetry:
  \[
  \mathcal{L}_{\text{int}} = \sum_i \xi_n \cdot \varphi \cdot \partial_\mu \gamma^i,
  \]
  where $\gamma^i$ are toroidally compactified internal symmetry coordinates.

  \item $\mathcal{L}_{\text{obs}}$: Observer-coupling and decoherence regulator:
  \[
  \mathcal{L}_{\text{obs}} = \lambda_O \cdot O^{\mu\nu} \nabla_\mu \phi \nabla_\nu \phi - \xi_n \cdot S_{\text{ent}}(\rho_{\text{red}}).
  \]

  \item $\mathcal{L}_{\text{ent}}$: Recursive entropy penalty:
  \[
  \mathcal{L}_{\text{ent}} = -\lambda_S \cdot S_{\text{rel}}(\rho_{\phi_n} \| \rho_{\phi_{n-1}}),
  \]
  where $S_{\text{rel}}$ is the quantum relative entropy.
\end{itemize}

\subsection*{B.2 Recursive Evolution and Filtering Constraint}

Recursive evolution of the configuration field is filtered through attractor-aligned updates:
\[
\phi_{n+1} = \phi_n \oplus \delta \phi_n,
\quad \text{subject to} \quad \delta \phi_n \in \text{Im}(\mathcal{P}_{G_{\text{coh}}}).
\]
The variation of the action is constrained accordingly:
\[
\delta \mathcal{A}_n[\phi] = 0 \quad \text{with} \quad \delta \phi_n \in \mathfrak{g}_{\text{coh}}.
\]
This ensures that the Euler–Lagrange equations reflect coherence-preserving dynamics only.

\subsection*{B.3 Attractor Stability and Potential Landscape}

The attractor \( \Psi^*(\phi) \) defines a fixed point of recursive evolution. Its stability is analyzed via the configuration-space Hessian:
\[
\mathcal{H}_{ij} = \frac{\partial^2 \mathcal{A}_n}{\partial \phi^i \partial \phi^j} \Bigg|_{\phi = \Psi^*(\phi)}.
\]
Lyapunov stability requires:
\[
\mathcal{H} \succeq 0 \quad \text{on} \quad \mathcal{P}_{G_{\text{coh}}}(\mathcal{C}_n),
\]
ensuring that evolution converges toward rather than diverges from memory-aligned attractor states.

\subsection*{B.4 Thermodynamic Enforcement}

All recursive transitions are regulated by entropy bounds and tension constraints:
\[
\mathcal{A}_n \text{ minimized under:} \quad 
\Delta S_{\text{fwd}} = \Delta S_{\text{mem}}, 
\quad \kappa_n \leq \kappa_C,
\]
with:
\[
\kappa_n = \frac{I(\phi_n, \phi_{n-1})}{\lambda_n}.
\]
This formalizes a dual thermodynamic–coherence constraint on evolution.

\subsection*{B.5 Summary}

Appendix B formalizes the Lagrangian components of recursive evolution, showing how gravitational, gauge, scalar, and observer fields are interwoven into a single attractor-constrained variational structure. All dynamics are filtered through entropy penalties, coherence conditions, and symmetry-aligned memory inheritance.
