\section*{Appendix A\\Recursive Symmetry and Attractor Constraints}
\addcontentsline{toc}{section}{Appendix A: Recursive Symmetry and Attractor Constraints}
\label{appendix:A}

\subsection*{A.1 Configuration Space and Symmetry Embedding}

We define the recursive configuration space as:
\[
\mathcal{C}_n = \mathbb{T}^6 = \text{Tor}(a_n) \times \text{Tor}(\varphi_n) \times \text{Tor}(\lambda_n) \times \text{Tor}(E_n) \times \text{Tor}(\gamma_n) \times \text{Tor}(\xi_n),
\]
where each degree of freedom evolves over a compact dimension with phase continuity. These toroidal embeddings allow recursive coherence and phase-filtered interference patterns across cycles without divergence.

Symmetry transformations are elements of a bounded transformation algebra \( \mathfrak{T} \), acting on \( \mathcal{C}_n \). Physical transitions are not arbitrary—they must preserve recursive alignment with the attractor \( \Psi^*(\phi) \). We define a coherence-preserving subset:
\[
G_{\text{coh}} \subset \text{Aut}(\mathcal{C}_n),
\]
where \( G_{\text{coh}} \) includes only those transformations that conserve entropy bounds, entanglement alignment, and observer resolution thresholds.

\subsection*{A.2 Attractor-Preserving Projection Operators}

We introduce a projection operator \( \mathcal{P}_{G_{\text{coh}}} \) that enforces attractor-preserving transformations:
\[
\mathcal{P}_{G_{\text{coh}}}[\delta \phi_n] := 
\begin{cases}
\delta \phi_n & \text{if } \delta \phi_n \in G_{\text{coh}}, \\
0 & \text{otherwise}.
\end{cases}
\]
This operator appears explicitly in the recursive action as a filter on legal transitions. It replaces arbitrary diffeomorphisms with attractor-conditioned pathways:
\[
\delta \mathcal{A}_n[\phi] = 0 \quad \text{subject to} \quad \delta \phi_n \in \text{Im}(\mathcal{P}_{G_{\text{coh}}}).
\]

\subsection*{A.3 Symmetry Filtering in the Transition Kernel}

The recursive kernel is defined by a coherence-weighted sum over attractor-aligned paths:
\[
K(\phi, \phi') = \sum_{\delta \phi \in G_{\text{coh}}} \mathcal{A}_{\delta \phi}(\phi, \phi') \cdot e^{-S_{\text{rel}}(\rho_\phi \| \rho_{\phi'})},
\]
where:
\begin{itemize}
  \item \( \mathcal{A}_{\delta \phi} \): amplitude for each transition,
  \item \( S_{\text{rel}} \): quantum relative entropy, penalizing misaligned transitions.
\end{itemize}

Transformations not in \( G_{\text{coh}} \) are exponentially suppressed:
\[
\mathcal{A}_{\delta \phi} \sim 0 \quad \text{if } \delta \phi \notin G_{\text{coh}}.
\]
This ensures recursive evolution flows only through the attractor’s stability basin.

\subsection*{A.4 Observer-Dependent Resolution Filtering}

Let \( \mathcal{P}_\xi \) denote a resolution-limited projection operator:
\[
\mathcal{P}_\xi[\rho] = \sum_{i \in \mathcal{B}_\xi} \ket{i}\bra{i} \rho \ket{i}\bra{i},
\]
where \( \mathcal{B}_\xi \) is a coarse-grained basis determined by observer resolution \( \xi \). This operator enforces decoherence filtering and constrains which components of the configuration space contribute to physical observables.

The recursive state update becomes:
\[
\rho \to \mathcal{P}_\xi \circ \mathcal{P}_{G_{\text{coh}}}[\rho],
\]
integrating both intrinsic symmetry and extrinsic observational constraints.

\subsection*{A.5 Summary}

This appendix formalizes how recursive evolution is shaped by coherence-preserving symmetry filtering in a compactified configuration space. Transformations are constrained by:
\begin{itemize}
  \item Alignment with the attractor \( \Psi^*(\phi) \),
  \item Entropy conservation and coherence fidelity,
  \item Observer-limited projection via resolution index \( \xi_n \).
\end{itemize}

This structure defines a recursive geometric logic for allowable transitions, enforcing that only memory-aligned, entropically admissible transformations survive the evolution across cosmological cycles.
