\section{Configuration Space and Symmetry Groups}

\subsection{Field Content}

We define an extended recursive configuration state vector
\[
\phi_n = \left( a_n, \varphi_n, \lambda_n, E_n, \gamma_n, \xi_n \right),
\]
where $(a_n, \varphi_n, \lambda_n, E_n)$ are inherited from the cosmological kernel formalism developed in \textit{The Self-Remembering Universe}~\cite{parian2025selfrememberingI}, and $(\gamma_n, \xi_n)$ introduce internal symmetry parameters and observer-defined resolution indices, respectively. The internal degrees of freedom $\gamma_n$ represent generalized gauge orientations, while $\xi_n$ denotes the local interpretive resolution of the observer, which affects state decoherence and measurement thresholds~\cite{zurek2003decoherence, nielsen2010quantum}.

\subsection{Dimensional Embedding and Compactification}

We assume a 12-dimensional recursive manifold $\mathcal{M}^{(12)}$, with compactified internal dimensions that encode memory structure and symmetry modulation rather than acting as purely geometric subspaces. This includes a toroidal submanifold representing internal field-space symmetry cycles. The recursive evolution of $\phi_n$ is governed by a non-Markovian transition kernel $K(\phi_{n}, \phi_{n-1})$~\cite{ashtekar2006quantum}, constrained by inherited memory and internal symmetry coherence.

\subsection{Gauge Symmetries and Coherence Filtering}

We define the recursive symmetry group as a coherence-preserving subgroup:
\[
G_{\text{coh}} \subset \mathrm{Aut}(\mathcal{C}_n),
\]
where $\mathcal{C}_n$ is the recursive configuration space at cycle $n$. The group $G_{\text{coh}}$ acts on $\phi_n$ via transformations that preserve attractor alignment and coherence fidelity. These operations are filtered by memory overlap constraints, entanglement eigenvalues, and observer-defined resolution thresholds.

Each configuration update $\phi_n \to \phi_{n+1}$ is permitted only if the associated transformation lies within $G_{\text{coh}}$, ensuring that evolution remains consistent with the recursive attractor $\Psi^*(\phi)$. This coherence filtering defines a selection rule: among all mathematically valid transitions, only those that preserve recursive memory and symmetry alignment are physically realized.

Observer input $\xi_n$ selects a resolution scale, effectively modulating the projection of internal states and regulating the degree of coherence preserved during transition. This mechanism embeds observational influence into the symmetry-constraint structure itself, enabling decoherence to shape physical evolution in a relational and memory-aware manner.
