\begin{abstract}
We propose a recursive quantum cosmological framework in which the universe evolves through coherence-filtered transitions between quantum geometries, preserving memory across bounces via entanglement-regulated interference. The model unifies three foundational principles: (1) loop quantum cosmology's discrete geometry, (2) non-Markovian decoherence governed by an entropy-sensitive memory kernel \( D(\tau, E) \), and (3) Einstein–Rosen bridge thermodynamics as the channel of recursive information transfer.

The configuration space \( \phi = (a, \varphi, \lambda, E) \) encodes the scale factor, scalar field amplitude, inter-cycle memory fidelity, and ERB entanglement eigenvalue. Evolution is governed by a transition kernel \( K(\phi, \phi') \), derived from spinfoam amplitudes and shaped by coherence filtering and entropy penalties. A total action \( \mathcal{A}_{\text{total}} \) integrates gravitational, informational, and decoherence contributions. A thermodynamic compensation constraint enforces entropy–tension balance: forward entropy gain must be offset by memory radiation and coherence strain. The recursive attractor \( \Psi^*(\phi) \) emerges as the unique fixed point of this evolution, with convergence guaranteed by a contraction mapping on Hilbert space.

The model yields falsifiable predictions, including: (i) low-\( \ell \) CMB suppression from phase-filtered memory overlap, (ii) quantized dips in the gravitational wave spectrum from recursive spin interference, (iii) scale-dependent non-Gaussianity modulated by inter-cycle fidelity, and (iv) void-aligned EB-mode polarization from entangled curvature domains. These signatures are testable by missions such as \textbf{LiteBIRD}, \textbf{CMB-S4}, \textbf{LISA}, and \textbf{Euclid}.

This framework provides a mathematically complete, memory-regulated alternative to inflationary and conformal cyclic cosmologies, positioning recursive coherence—rather than initial conditions—as the fundamental organizing principle of cosmic evolution.
\end{abstract}
