\section{Attractor Dynamics, Recursive Interference, and Entropy Flow}
\label{sec:recursive-variational}

The evolution of the universe across cycles is governed not only by local dynamics, but by quantum interference between configuration states. These transitions are constrained by entanglement structure, entropy divergence, and recursive memory fidelity. We formalize this using a recursive action principle and define a stable attractor \( \Psi^*(\phi) \) governing long-term coherence.

\subsection{Recursive Configuration and Action}
\label{subsec:recursive-action}

We define the full cosmological configuration as:
\[
\phi = (a, \varphi, \lambda, E),
\]
where \( a \) is the scale factor, \( \varphi \) the scalar field value, \( \lambda \) the cycle-to-cycle coherence fidelity, and \( E = \sqrt{-\mathrm{Tr}(\rho_{\text{red}} \log \rho_{\text{red}})} \) the entanglement eigenvalue across the Einstein–Rosen bridge.

The recursive action governing evolution is:
\begin{equation}
\mathcal{A}_n = \sum_{k=1}^n \left[ \mathcal{A}_{\text{EH}}[\phi_k] + \lambda_E D(\tau_k, E_k) - \gamma \, S(\rho_{\phi_k} \| \rho_{\phi_{k-1}}) \right],
\end{equation}
as detailed in Appendix~\ref{appendix:C6}, where:
\begin{itemize}
    \item \( \mathcal{A}_{\text{EH}} \): Einstein–Hilbert gravitational action per cycle,
    \item \( D(\tau_k, E_k) \): non-Markovian memory kernel weighted by entanglement,
    \item \( S(\rho_k \| \rho_{k-1}) \): quantum relative entropy measuring information loss.
\end{itemize}

\subsection{Recursive Interference and Attractor Definition}
\label{subsec:attractor}

State evolution follows the kernel-mediated recursion:
\[
\Psi_{n+1}(\phi) = \int K_{\text{norm}}(\phi, \phi') \Psi_n(\phi') \, d\phi',
\]
where \( K_{\text{norm}} \) is the normalized transition kernel (see Section~\ref{sec:kernel}, Eq.~4.1). The attractor \( \Psi^*(\phi) \) is defined as the unique fixed point:
\[
\Psi^*(\phi) = \int K_{\text{norm}}(\phi, \phi') \Psi^*(\phi') \, d\phi', \quad \text{with} \quad \int |\Psi^*(\phi)|^2 d\phi = 1.
\]
Convergence toward the attractor is confirmed when:
\[
\| \Psi_{n+1} - \Psi_n \|^2 < \epsilon, \quad \text{for some} \quad \epsilon \ll 1.
\]

The attractor also minimizes a coherence-weighted entropy functional:
\[
\Psi^* = \arg\min_{\Psi} \left\{ \lambda_S S(\rho_\Psi) + \lambda_T \lambda^2 - \lambda_C |\langle \Psi | \Psi_{n-1} \rangle|^2 \right\},
\]
subject to normalization \( \langle \Psi | \Psi \rangle = 1 \), as derived in Appendix~\ref{appendix:C7}. This structure balances entropy, memory strain, and fidelity with past cycles.

\subsection{Entropy–Tension Compensation and Collapse Thresholds}
\label{subsec:entropy-dynamics}

Recursive evolution obeys a thermodynamic constraint equating entropy flow and coherence tension:
\[
\frac{d}{dn} \left( S_n + \lambda_n \right) = \frac{dS_{\text{rad}}}{dn},
\]
where entropy loss from coherence sharpening is compensated by radiation. This enforces thermodynamic consistency across cycles and is derived in Appendix~\ref{appendix:C8}.

When coherence tension exceeds a critical threshold \( \lambda_n > \lambda_{\text{crit}} \), recursive failure occurs, interpreted as a rupture event akin to a supernova-like collapse or dimensional reset:
\[
\lambda_n > \lambda_{\text{crit}} \quad \Rightarrow \quad \text{GW burst at } f_{\text{burst}} \sim \tau_{\text{mem}}^{-1}.
\]

\subsection{Observable Consequences of the Attractor}
\label{subsec:observables}

The attractor structure and filtering kernel yield several falsifiable predictions:

\paragraph{CMB Power Suppression:}
\[
C_\ell^{\text{rec}} \propto \exp\left[-\frac{(\varphi - \varphi')^2}{2\sigma_\varphi^2}\right] \cdot \cos^2(\Delta \theta),
\]
where \( \sigma_\varphi \) is the coherence filter width and \( \Delta \theta \) is the recursive interference phase. This predicts low-\( \ell \) suppression and angular correlations.

\paragraph{Gravitational Wave Echoes:}
\[
f_j \sim \frac{1}{\tau_M} \cdot e^{-j \Delta}, \quad j \in \mathbb{Z}^+,
\]
from boundary-induced memory delays and entanglement reemission across ERBs.

\paragraph{Entropy Bound Enforcement:}
\[
S(\rho_\phi \| \rho_{\phi'}) < \frac{A(\phi)}{4G} \quad \text{if } E > E_{\text{crit}},
\]
enforcing bounded information divergence when entanglement exceeds the critical propagation threshold.

\subsection{Summary}

This section establishes the recursive attractor \( \Psi^*(\phi) \) as a stable, self-consistent eigenstate of the transition kernel. It emerges from a coherence-filtered entropy variational principle and governs the memory dynamics of the universe. Recursive convergence is guaranteed by contraction mapping, while tension–entropy dynamics enforce radiative compensation and yield falsifiable observables in gravitational waves, CMB structure, and information entropy bounds.
