\section{Recursive Dynamics and the Fixed-Point Attractor}
\label{sec:attractor}

\subsection{10.1 Recursive Evolution and Attractor Definition}

We define the recursive wavefunction \( \Psi_n(\phi) \) as the quantum configuration at cycle \( n \). Recursive evolution is governed by the normalized transition kernel \( K_{\text{norm}}(\phi, \phi') \):
\[
\Psi_{n+1}(\phi) = \int K_{\text{norm}}(\phi, \phi') \Psi_n(\phi') \, d\phi',
\]
where \( K_{\text{norm}} \) is defined in Section~\ref{sec:kernel} and formally derived in Appendix~\ref{appendix:C9}.

The recursive attractor \( \Psi^*(\phi) \) is the unique fixed point of this operator:
\[
\Psi^*(\phi) = \int K_{\text{norm}}(\phi, \phi') \Psi^*(\phi') \, d\phi',
\]
representing a coherence-preserving, entropy-stabilized eigenstate of the full recursive system. Its existence and convergence are mathematically proven via contraction mapping (Appendix~\ref{appendix:C9}).

\subsection{10.2 Conditions for Convergence}

Convergence to \( \Psi^*(\phi) \) is driven by recursive filtering and entropy suppression. The key conditions are:

\begin{enumerate}
    \item \textbf{Entropy Bound:}
    \[
    S_n = -\text{Tr}[\rho_n \ln \rho_n] < S_{\text{max}}(E, n)
    \]
    \item \textbf{Memory Fidelity:}
    \[
    M_n = |\langle \Psi_n | \Psi_{n-1} \rangle|^2 > M_{\text{crit}}
    \]
    \item \textbf{Fitness Threshold:}
    \[
    \mathcal{F}_n = \alpha_C \cdot \text{Tr}(\rho_n^2) - \alpha_S S_n + \alpha_M M_n \geq \mathcal{F}_{\text{crit}}
    \]
\end{enumerate}

The attractor also minimizes a coherence-weighted entropy functional (Appendix~\ref{appendix:C7}):
\[
\Psi^* = \arg\min_{\Psi} \left\{ \lambda_S S(\rho_\Psi) + \lambda_T \lambda^2 - \lambda_C |\langle \Psi | \Psi_{n-1} \rangle|^2 \right\}.
\]

\subsection{10.3 Dynamical Properties of \( \Psi^*(\phi) \)}

The attractor exhibits the following properties:

\begin{itemize}
    \item \textbf{Lyapunov Stability:}
    \[
    \mathcal{L}_n = \log \left| \frac{\|\Psi_{n+1} - \Psi_n\|}{\|\Psi_n - \Psi_{n-1}\|} \right| < 0
    \]
    \item \textbf{Entropy Saturation:}
    \[
    S(\rho_{n+1} \| \rho_n) \to 0
    \]
    \item \textbf{Phase Coherence:}
    \[
    \Delta \theta_n \to \theta^* \in [0, \pi]
    \]
\end{itemize}

These behaviors signal long-term stability of the recursive system under entropy filtering and memory preservation.

\subsection{10.4 Physical Interpretation}

The attractor \( \Psi^*(\phi) \) defines the equilibrium limit of recursive quantum evolution. It acts as:
\begin{itemize}
    \item A \textbf{memory-retaining quantum state} stable under interference filtering,
    \item A \textbf{conformal-invariant configuration}:
    \[
    \Psi^*(\Omega^2 g_{\mu\nu}, \phi) = \Psi^*(g_{\mu\nu}, \phi),
    \]
    \item A \textbf{selective filter}: configurations that interfere constructively with \( \Psi^* \) persist; others decohere.
\end{itemize}

\subsection{10.5 Observational Implications}

Attractor convergence yields testable signatures:

\begin{itemize}
    \item \textbf{CMB Suppression:}
    \[
    C_\ell^{\text{rec}} \propto \exp\left[-\frac{(\varphi - \varphi')^2}{2\sigma_\varphi^2}\right] \cdot \cos^2(\Delta \theta)
    \]
    \item \textbf{Recursive Non-Gaussianity:}
    \[
    f_{\text{NL}}^{\text{rec}} \propto \lambda_E |\langle \Psi_{n-1} | \Psi_n \rangle|^\alpha
    \]
    \item \textbf{Gravitational Wave Echoes:}
    Phase-coherent echo modes cluster around recursive intervals.
\end{itemize}

\subsection{10.6 Recursive Entropy Compensation and Breakdown}

Recursive evolution imposes a thermodynamic constraint on coherence tension \( \lambda_n \). The entropy-tension balance condition (Appendix~\ref{appendix:C8}) reads:
\[
\Delta S_{\text{gain}} + \Delta S_{\text{rad}} = \Delta S_{\text{exp}},
\]
ensuring that entropy loss from increased memory (gain) is offset by Hawking radiation (rad) and dilution from expansion (exp).

A more direct constraint relates bounce entropy and radiative emission:
\[
\frac{d\lambda_n}{dn} + \frac{dS_{\text{rad}}^{(n)}}{dn} = \frac{dS_{\text{bounce}}^{(n)}}{dn},
\]
where:
\begin{itemize}
    \item \( S_{\text{bounce}}^{(n)} = \frac{A_n}{4G\hbar} \),
    \item \( S_{\text{rad}}^{(n)} \): entropy emitted via Hawking radiation across the ERB.
\end{itemize}

When:
\[
\lambda_n > \lambda_{\text{crit}},
\]
the system undergoes recursive collapse or string rupture. The severity depends on the number of dimensions whose tension threshold is breached, corresponding to distinct classes of supernovae or memory-driven cosmic events.

\subsection{10.7 Summary}

The attractor state \( \Psi^*(\phi) \) encodes the long-term structure of recursive memory, defined by stability under entropy-filtered interference and convergence under contraction. It is both a mathematical fixed point and a physical coherence filter: the universe retains only configurations consistent with recursive survival. Its dynamics are falsifiable, predictive, and central to the model’s explanatory power.
