\section*{Appendix B\\Single-Cycle Decoherence and Memory Kernel\\Dynamics}
\addcontentsline{toc}{section}{Appendix B: Single-Cycle Decoherence and Memory Kernel Dynamics}
\label{appendix:B}

\subsection*{Overview}

Intra-cycle decoherence is modeled using a density matrix formalism \( \rho_n(t) \), which evolves under a non-Markovian master equation coupled to memory feedback. Each cycle begins with a configuration-weighted projection from the previous epoch, initialized through the transition kernel \( K(\phi, \phi') \) (defined in Appendix~\ref{appendix:C2}). The model captures how recursive coherence propagates within a single cycle before subsequent kernel updates.

\subsection*{Recursive Kernel Initialization}

The initial state of cycle \( n \) is defined as:
\[
\rho_n(\phi) = \int K(\phi, \phi') \, \rho_{n-1}(\phi') \, d\phi'
\]
with kernel structure:
\[
K(\phi, \phi') \sim \exp\left[i S_{\text{ERB}}(\phi, \phi')\right] \cdot \mathcal{F}(\phi, \phi')
\]
The filter:
\[
\mathcal{F}(\phi, \phi') = \exp\left[
-\frac{(a - a')^2}{2\sigma_a^2}
-\frac{(\varphi - \varphi')^2}{2\sigma_\varphi^2}
-\frac{(E - E')^2}{2\sigma_E^2}
\right]
\]
selects aligned configurations in geometry, field amplitude, and entanglement energy. The phase term \( S_{\text{ERB}} \) is the entropic bridge action (Appendix~C.3).

\subsection*{Non-Markovian Decoherence Dynamics}

The system evolves via:
\[
\dot{\rho}_n(t) = -i[H, \rho_n(t)] + \int_0^t D(\tau) \, [\hat{O}, [\hat{O}, \rho_n(t - \tau)]] \, d\tau
\]
where:
\begin{itemize}
    \item \( H \): Effective LQC Hamiltonian,
    \item \( \hat{O} = \mathrm{Tr}(\rho_n^2) \, \hat{R}_{\text{eff}}(x) \): Purity-weighted curvature operator,
    \item \( \hat{R}_{\text{eff}} \): Effective scalar curvature fluctuation operator modulated by local entanglement structure.
\end{itemize}
This form captures decoherence induced by gravitational degrees of freedom within each minisuperspace slice.

\subsection*{Entropy-Modulated Memory Kernel}

The kernel \( D(\tau) \) evolves dynamically based on the entropy:
\[
S_n(t) = -\mathrm{Tr}[\rho_n(t) \log \rho_n(t)]
\]
We define:
\[
D(\tau) = \gamma(S_n) \, e^{-\tau/\tau_c(S_n)} \cos(\omega_0(S_n) \tau)
\]
with entropy-dependent parameters:
\begin{align*}
\tau_c(S_n) &\sim S_n^{-1} \quad \text{(coherence time)} \\
\omega_0(S_n) &= \omega_{\text{LQG}} \sqrt{1 - \left(S_n / S_{\text{max}}\right)} \\
\gamma(S_n) &\propto e^{-\beta S_n}
\end{align*}
Here, \( S_{\text{max}} \) is a cycle-specific entropy bound derived from geometric constraints (e.g., ERB area). This form ensures:
\begin{itemize}
  \item Markovian behavior as \( S_n \to 0 \),
  \item Coherence breakdown at saturation \( S_n = S_{\text{max}} \),
  \item Spectral coherence tied to spin-quantized LQG structure.
\end{itemize}

\subsection*{Numerical Priorities and Observables}

Key computational targets for each cycle \( n \):

\paragraph{Entropy Evolution:}
Track the decoherence process via:
\[
S_n(t) = -\mathrm{Tr}[\rho_n(t) \log \rho_n(t)]
\]

\paragraph{Intra-cycle Fidelity:}
Evaluate memory retention:
\[
\mathcal{F}_n(t) = \mathrm{Tr}[\rho_n(0)\rho_n(t)]
\]
This differs from the inter-cycle fidelity \( \lambda_n = |\langle \Psi_{n-1} | \Psi_n \rangle|^2 \) used in attractor convergence (Appendix~C.1).

\paragraph{Kernel Calibration:}
Tune the entropy-dependent functions \( \gamma(S_n) \), \( \tau_c(S_n) \), and \( \omega_0(S_n) \) to fit:
\begin{itemize}
  \item Low-\( \ell \) CMB power suppression,
  \item EB-mode polarization structure,
  \item Coherence filtering thresholds near LISA-scale gravitational wave bands.
\end{itemize}

\paragraph{Initial Condition:}
Each cycle begins from:
\[
\rho_n(0) = \int K(\phi, \phi') \, \rho_{n-1}(\phi') \, d\phi'
\]

Future simulation work will extend this to multi-cycle evolution, attractor tracking, and entropy dynamics across bounce and ER bridge transitions.
