\section*{Appendix D\\Waveform Collapse and the Relativistic Geometry of Light-Speed Limits}
\addcontentsline{toc}{section}{Appendix D: Waveform Collapse and the Relativistic Geometry of Light-Speed Limits}
\label{appendix:D}

We propose a geometric and informational reinterpretation of the relativistic speed limit as a coherence-collapse threshold, governed by the internal waveform dynamics of an observer. Within the recursive cosmology framework, acceleration is interpreted as a deformation of the observer’s internal spacetime waveform. The speed of light \( c \) marks the critical point beyond which this waveform collapses, severing entanglement and halting recursive memory propagation.

\subsection*{D.1 Spacetime as a Sinusoidal Carrier Wave}

Let the internal structure of an observer be represented by a coherence-preserving waveform:
\[
\Psi(x,t) = A \sin(kx - \omega t)
\]
where \( A \) is the amplitude, \( k \) is the spatial wavenumber, and \( \omega \) is the frequency. Lorentz boosts induce compression:
\[
x' = \gamma(x - vt), \quad t' = \gamma\left(t - \frac{vx}{c^2}\right)
\]
causing blue-shift and waveform squeezing.

As \( v \to c \), the waveform collapses to a singular peak:
\[
\lim_{v \to c} \Psi(x,t) \to \delta(x - ct)
\]
This collapse eliminates phase structure and coherence bandwidth, marking the loss of information propagation.

\subsection*{D.2 Recursive Collapse and Tension Threshold}

In the recursive framework, coherence propagation is governed by the memory fidelity \( \lambda_n \), and entanglement eigenvalue \( E \). We introduce a cycle-specific tension constraint:
\[
\lambda_n := \text{maximum recursive string tension}
\]
where high tension encodes high coherence bandwidth and low entropy. As velocity increases, internal tension increases with information gain. The limit \( \lambda_n \to \lambda_{\text{max}} \) defines a coherence singularity.

Waveform collapse is therefore associated with:
\[
\frac{d\lambda_n}{dv} > 0, \quad \lim_{v \to c} \lambda_n \to \infty
\]
At the point of divergence, the string “snaps,” resulting in coherence rupture and initiating a reset event—interpreted cosmologically as a supernova or bounce.

\subsection*{D.3 Thermodynamic Balance and Entropy Drift}

This collapse corresponds to a failure in the entropy–memory balance:
\[
\Delta S_{\text{fwd}} > \Delta S_{\text{mem}} \Rightarrow \text{Collapse}
\]
Hawking radiation is required to restore equilibrium. The thermodynamic equation governing this behavior becomes:
\[
\frac{dS}{dt} \approx -\frac{d\lambda_n}{dt} + \Phi_H
\]
where \( \Phi_H \) denotes Hawking flux. The system must radiate away sufficient information to reduce recursive tension below the collapse threshold. Failure leads to entropic overload and decoherence of the memory kernel:
\[
D(\tau, E) \to 0, \quad \lambda_n \to 0
\]

\subsection*{D.4 Collapse Signature in the Kernel}

Collapse manifests geometrically in the recursive kernel:
\[
K(\phi, \phi') \to \delta(\phi - \phi'), \quad \mathcal{F}(\phi, \phi') \to 0
\]
indicating loss of interference bandwidth. Recursive propagation halts, and a new cycle is initialized with minimal inherited structure:
\[
\Psi_{n+1}(\phi) \sim \Psi_{\text{vac}}(\phi)
\]

\subsection*{D.5 Causal Geometry and Spin Network Collapse}

In loop quantum gravity, boost-induced tension compresses spin networks:
\begin{itemize}
  \item Spatial areas shrink: \( A_f \to \ell_{\text{Pl}}^2 \)
  \item Node connectivity vanishes: \( \iota_v \to 0 \)
  \item Information channels sever across ERB boundaries
\end{itemize}
This is consistent with the interpretation of relativistic collapse as a transition to a disconnected quantum geometry—coherence fails, and recursion terminates.

\subsection*{D.6 Falsifiable Prediction}

A falsifiable implication of this model is that any physical system approaching the relativistic coherence limit must exhibit observable loss of recursive fidelity, manifested as:
\[
\lambda_n(f_{\text{GW}}) \to 0 \quad \text{as} \quad f \to f_c
\]
for a critical frequency \( f_c \sim \lambda_{\text{max}}^{-1/2} \). This predicts sharp dips in the gravitational wave spectrum at tension-induced collapse thresholds, distinguishable from inflationary noise floors.

\subsection*{D.7 Summary}

We reinterpret the relativistic limit as a coherence-collapse threshold enforced by recursive string tension \( \lambda_n \), entropy bounds, and memory kernel viability. The light-speed barrier marks a causal-holographic boundary beyond which entangled identity and memory propagation cannot persist. This offers a thermodynamically and geometrically grounded explanation for the structure of relativity within a recursive cosmological system.
